### 2016-09-19 paper/paper101.tex
created folder for paper

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Earvin Balderama\\Loyola University Chicago \And 
        Taylor Trippe\\Loyola University Chicago}
\title{\pkg{hurdlr}: An \proglang{R} Package for Zero-inflated and Extremely Over-dispersed Count Data}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Earvin Balderama, Taylor Trippe} %% comma-separated
\Plaintitle{hurdlr: An R Package for Zero-inflated and Extremely Over-dispersed Count Data} %% without formatting
\Shorttitle{\pkg{hurdlr}: An R Package for Count Data} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  The abstract of the article.
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Earvin Balderama\\
  Department of Mathematics and Statistics\\
  Loyola University Chicago\\
  1032 W Sheridan Rd., \\
  Chicago, IL 60660, USA\\
  E-mail: \email{ebalderama@luc.edu}\\
  URL: \url{http://math.luc.edu/~ebalderama/}\\

  Taylor Trippe\\
  Department of Mathematics and Statistics\\
  Loyola University Chicago\\
  1032 W Sheridan Rd., \\
  Chicago, IL 60660, USA\\
  E-mail: \email{ebalderama@luc.edu}\\
  URL: \url{http://math.luc.edu/~ebalderama/}
}

%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\section{Introduction}

\section{Models for count data}

\section{Double-hurdle models}

\section[Implementing hurdlr]{Implementing \pkg{hurdlr}}

\section{Discussion}

\section*{References}

\end{document}